Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2024-08-08 paper/main.tex
Updates
updates

In this section, we develop a novel measure of partial association for mixed
data by extending partial correlations \citep{Whittaker2009} \todo{Verify this
citation} to mixed data. To compute partial correlations we use: 1) A
residualization method, and 2) a measure of associations between the residuals.
In the case when both $ X $ and $ Y $ are continuous variables, the residuals
are typically computed using a regression based method and  Pearson's
correlation coefficient is used as a measure of association between the
residuals. To extend this idea to mixed data, we combining a mixed data
residualization method \citep{Ankan2023} with a multivariate measure of
association based, Pillai's Trace \citep{Pillai1955}.

Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $, $ Y $, and $
\bm{Z} $, we are interested in computing an estimate of the conditional
association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $ conditioned on $
\bm{Z} $. To compute the residuals, we utilize the residualization approach
from \citet{Ankan2023}. Depending on the type of dependent variable, the
residuals are defined as follows:

@@

\begin{equation*}
	\begin{split}
		\bm{\rho}(\bm{U}, \bm{V}) &= \sqrt{\mathrm{E} \left( \Sigma_{\bm{UU}}^{-1} \Sigma_{\bm{UV}} \Sigma_{\bm{VV}}^{-1} \Sigma_{\bm{VU}} \right)}
	\end{split}
\end{equation*}

@@

Many measures of association based on canonical correlation have been used in
multivariate statistics with each of them having slightly different properties.
\begin{itemize}
	\item Wilks' Lambda: $ \Lambda = \prod (1 - \phi_i^2) $
	\item Roy's Largest Root: $ \theta = \phi_{\max}^2 $
	\item Pillai's Trace: $ \tau = \sum \phi_i^2 $
\end{itemize}

@@

$$ \rho(X, Y; \bm{Z}) = \sum_{i=1}^{min(\rvert R_X \rvert, \rvert R_Y \rvert)} (\phi_i(R_X, R_Y))^2 $$

@@

\begin{enumerate}
	\item Bounded between $ 0 $ and $ 1 $.
	\item Independent of sample size.
	\item Invariant to the one-hot encoding scheme used fr categorical variables.
	\item Reduces to partial correlations when both variables are continuous.
	\item In case of discrete variables, connected to Cramer\'s V, a well
		known effect size measure for discrete variables.
\end{enumerate}
\todo[inline]{Expand on these points}

>>>>

In this section, we introduce a novel measure of partial association for mixed
data by extending the concept of partial correlations \citep{Whittaker2009}
\todo{Verify this citation} to handle mixed data types. To compute partial
correlations, we have two key steps: 1) a residualization method, and 2) a
measure of associations between the residuals. When both $ X $ and $ Y $ are
continuous variables, residuals are typically computed using a regression-based
method, and Pearson's correlation coefficient is used to measure the association
between these residuals. To extend this approach to mixed data, we combine a mixed
data residualization method \citep{Ankan2023} with a multivariate measure of
association based on canonical correlations, Pillai's Trace \citep{Pillai1955}.

Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $, $ Y $, and $
\bm{Z} $, we aim to estimate the conditional
association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $ conditioned on $
\bm{Z} $. We use the residualization approach from \citet{Ankan2023} to compute 
the residuals. The residuals are defined as follows, depending on the type of 
dependent variable:

@@

\begin{equation}
	\begin{split}
		\bm{\rho}(\bm{U}, \bm{V}) &= \sqrt{\mathrm{E} \left( \Sigma_{\bm{UU}}^{-1} \Sigma_{\bm{UV}} \Sigma_{\bm{VV}}^{-1} \Sigma_{\bm{VU}} \right)}
	\end{split}
\end{equation}

@@

Many measures of association based on canonical correlation have been used in
multivariate statistics with each of them having slightly different properties.
\begin{itemize}
	\item Wilks' Lambda:
		$$ \Lambda = \prod (1 - \phi_i^2) $$
	\item Roy's Largest Root: 
		$$ \theta = \phi_{\max}^2 $$
	\item Pillai's Trace: 
		$$ \tau = \sum \phi_i^2 $$
\end{itemize}

@@

\begin{equation}
	\rho(X, Y; \bm{Z}) = \frac{1}{\min(\rvert R_X \rvert, \rvert R_Y \rvert)}
	\sum_{i=1}^{\min(\rvert R_X \rvert, \rvert R_Y \rvert)} (\phi_i(R_X, R_Y))^2
\end{equation}

@@

\begin{enumerate}
	\item \textbf{Bounded: } $ \rho $ is bounded between $ 0 $ and $ 1 $,
		with $ 0 $ representing no association while $ 1 $ representing
		perfect linear association. This helps in interpretation and
		comparing values across multiple partial associations.
	\item \textbf{Independent of sample size: }
	\item \textbf{Invariant to the dummy encoding scheme: } As we use dummy
		encoding to compute the residuals for categorical variables,
		but as canonical correlations search for linear combinations
		over the columns, the measure is invariant to the dummy
		encoding scheme that we use.
	\item \textbf{Generalizes partial correlations to mixed data: }
		As canonical correlation is equivalent to Pearson's correlation
		coefficient, when both $ X $ and $ Y $ are continuous, $ \rho $
		is equivalent to canonical correlations. Hence, $ \rho $ is a
		generalization of partial correlations to mixed data.
	\item \textbf{Related to Cramer\'s V: }
\end{enumerate}
\todo[inline]{Expand on these points}
